---
title: |
     | ST588 Homework 6
     | Matt Bray
format: html
code-overflow: wrap
editor: visual
---

```{r setup, include=FALSE}
#This code prevents warnings from displaying in rendered output.
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

*First, call packages that will be used in this assignment.*

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
```

## Task 1: Conceptual Questions

1.What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

:   -   `lapply()` applies a vectorized function to a list. The `map()` function is the `purrr` equivalent.

**2.Suppose we have a list called `my_list`. Each element of the list is a numeric data frame (all columns are numeric). We want to use `lapply()` to run the code `cor(numeric_matric, method = "kendall")` on each element of the list. Write code to do this below! (I'm really trying to ask you how you specify `method = "kendall"` when calling `lapply()`)**

```{r}
#this will create three data frames and combine them to create a list of length 3, "my_list" to test my answer.
A <- as_tibble(rnorm(100)) |>
  rename("a" = "value")|>
  mutate(b = rnorm(100)) |>
  mutate(c = rnorm(100))
  
B <- as_tibble(rnorm(100)) |>
  rename("a" = "value")|>
  mutate(b = rnorm(100)) |>
  mutate(c = rnorm(100))

C<- as_tibble(rnorm(100)) |>
  rename("a" = "value")|>
  mutate(b = rnorm(100)) |>
  mutate(c = rnorm(100))

A.df <- as.data.frame(A)
B.df <- as.data.frame(B)
C.df <- as.data.frame(C)

my_list <- list(A.df, B.df, C.df)

#This code applies the `cor` function to all elements of my_list.
lapply(X=my_list, FUN=cor, method="kendall")

```

3.What are two advantages for using `purrr` functions instead of the `BaseR` apply family?

:   -   There is more consistency between the functions' syntax
    -   And the helper functions

4.What is a side-effect function?

:   -   It is a function that returns something outside of the input dataset.
    -   For instance, `print()` would be a side-effect function, since it doesn't do anything to the original dataset.

5.Why can you name a variable `sd` in a function and not cause any issues with the `sd` function?

:   -   The custom function that contains `sd` has its own environment, and once the custom function is done execution, that environment ceases to exist.

## Task 2 - Writing R Functions

1.  Write a basic function (call it `getRMSE()`) tht takes in a *vector* of responses and a *vector* of predictions and outputs RMSE.

RMSE:

$$RMSE = \sqrt{\frac{1}{n}\sum_{i=1}^n (y_i - \hat{y}_i)^2}$$

-   If a value is missing for the vector of responses (i.e. an `NA` is present), allow for additional arguments to the `mean()`function (elipses) that removes the `NA` values in the computation.

*Create function `getRMSE()`:*

```{r}
getRMSE <- function(responses, predictions, ...){
  SE <- as.numeric(map2(responses, predictions, \(x,y) (x-y)^2))
  MSE <- mean(SE,...)
  RMSE <- sqrt(MSE)
  return(RMSE)
}
```

2.  The following code will generate some response values and predictions:

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

*Test `getRMSE()` function:*

```{r}
getRMSE(resp, pred)
```

*Add missing values and test `getRMSE()` function:*

```{r}
resp.1 <- resp
resp.1[13] <- NA_real_
resp.1[43] <- NA_real_
resp.1
```

*Now, we'll test the behavior of the `getRMSE` function without specifying behavior to deal with missing values...*

```{r}
getRMSE(resp.1, pred)
```

*... and **with** specifying behavior to deal with missing values*

```{r}
getRMSE(resp.1, pred, na.rm=TRUE)
```

3.  Create `getMAE` function for evaluating predictions:

$$MAE = \frac{1}{n}\sum_{i=1}^n |y_i - \hat{y}_i|$$

```{r}
getMAE <- function(responses, predictions, ...){
  AE <- as.numeric(map2(responses, predictions, \(x,y) abs(x-y)))
  MAE <- mean(AE,...)
  return(MAE)
}
```

*We'll test the `getMAE` function behavior without `na.rm`...*
```{r}
getMAE(resp.1, pred)
```
*...And then we'll test the `getMAE` function with the `na.rm` function:*
```{r}
getMAE(resp.1, pred, na.rm=TRUE)
```
*Here, we'll create a wrapper function that implements both `getRMSE` and `getMAE`.*
```{r}
evalPred <- function(responses, predictions, RMSE=TRUE, MAE=TRUE, ...){
  
  if(!is.numeric(responses)|!is.numeric(predictions)){
    stop("Input should be numeric (atomic) vector")
  } else if(!is.atomic(responses)|!is.atomic(predictions)){
    stop("Needs to be simpler data")
  } else if(!is.vector(responses)|!is.vector(predictions)){
    stop("this is not the vector you are looking for")
  }
  
  
  if(RMSE==TRUE){ 
    RMSE <- getRMSE(responses, predictions, ...)
  } else {
    RMSE <- as.numeric(NA)
  }
  
  if(MAE==TRUE){
    MAE <- getMAE(responses, predictions, ...)
  } else {
    MAE <- as.numeric(NA)
  }
  
  #output <- (RMSE, MAE)
  outRMSE <-as_tibble(RMSE) |>
    rename("RMSE"="value")
  
  outMAE <-as_tibble(MAE) |>
    rename("MAE" = "value")
  
  out <-merge(outRMSE, outMAE)
  
return(out)
}
```


Here, we'll test `evalPred` across multiple scenarios:

*First, we'll look at `RMSE` only:
```{r}
evalPred(resp, pred, MAE=FALSE)
```
Then, we'll look at `MAE` only:
```{r}
evalPred(resp, pred, RMSE=FALSE)
```
Then...both metrics:
```{r}
evalPred(resp, pred)
```

Test with missing data:
```{r}
evalPred(resp.1, pred, na.rm=TRUE)
```

Test with incorrect inputs:
```{r}
#try() allows for evalution and demonstrates error without stopping entire document render.
try(evalPred(resp, A.df))
```


## Task 3: Querying an API and a Tidy-Style Function

1. Use `GET()` from the `httr` package to return information about a topic that you are interested in that has been in the news lately, and store as R object.
```{r}
URL_ids<- "https://newsapi.org/v2/everything?q=spacex&apiKey=2a263eef22a94f42ab1bbd0802661e28"

id_info <- httr::GET(URL_ids)
str(id_info, max.level=1)

parsed <- fromJSON(rawToChar(id_info$content))
spacex_info <- as_tibble(parsed$articles)
spacex_info
x <- pluck(spacex_info)
x
```

