---
title: |
     | ST588 Homework 6
     | Matt Bray
format: html
code-overflow: wrap
editor: visual
---

*First, call packages that will be used in this assignment.*

```{r}
library(tidyverse)
```

## Task 1: Conceptual Questions

1.What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

:   -   `lapply()` applies a vectorized function to a list. The `map()` function is the `purrr` equivalent.

**2.Suppose we have a list called `my_list`. Each element of the list is a numeric data frame (all columns are numeric). We want to use `lapply()` to run the code `cor(numeric_matric, method = "kendall")` on each element of the list. Write code to do this below! (I'm really trying to ask you how you specify `method = "kendall"` when calling `lapply()`)**

```{r}
#this will create three data frames and combine them to create a list of length 3, "my_list" to test my answer.
A <- as_tibble(rnorm(100)) |>
  rename("a" = "value")|>
  mutate(b = rnorm(100)) |>
  mutate(c = rnorm(100))
  
B <- as_tibble(rnorm(100)) |>
  rename("a" = "value")|>
  mutate(b = rnorm(100)) |>
  mutate(c = rnorm(100))

C<- as_tibble(rnorm(100)) |>
  rename("a" = "value")|>
  mutate(b = rnorm(100)) |>
  mutate(c = rnorm(100))

A.df <- as.data.frame(A)
B.df <- as.data.frame(B)
C.df <- as.data.frame(C)

my_list <- list(A.df, B.df, C.df)

#This code applies the `cor` function to all elements of my_list.
lapply(X=my_list, FUN=cor, method="kendall")

```

3.What are two advantages for using `purrr` functions instead of the `BaseR` apply family?

:   -   There is more consistency between the functions' syntax
    -   And the helper functions

4.What is a side-effect function?

:   -   It is a function that returns something outside of the input dataset.
    -   For instance, `print()` would be a side-effect function, since it doesn't do anything to the original dataset.

5.Why can you name a variable `sd` in a function and not cause any issues with the `sd` function?

:   -   The custom function that contains `sd` has its own environment, and once the custom function is done execution, that environment ceases to exist.
